---
title: "NYC Flights 2013 Analysis"
format:
  html:
    code-fold: true
    warning: false
    message: false
  md:
    warning: false
    message: false
execute:
  freeze: false
---



```{r}
#| label: load-packages
#| echo: false
library(ggplot2)
library(nycflights13)
```

# Variable 1: Departure Delay

```{r}
#| label: plot-dep-delay
#| fig-cap: "Distribution of Departure Delays"
ggplot(nycflights13::flights, aes(x = dep_delay)) +
  geom_histogram(binwidth = 10, fill = "steelblue", color = "black") +
  xlim(-50, 300) +
  labs(title = "Departure Delays", x = "Delay (minutes)", y = "Count") +
  theme_minimal()
```

# Variable 2: Arrival Delay

```{r}
#| label: plot-arr-delay
#| fig-cap: "Distribution of Arrival Delays"
ggplot(nycflights13::flights, aes(x = arr_delay)) +
  geom_histogram(binwidth = 10, fill = "darkorange", color = "black") +
  xlim(-50, 300) +
  labs(title = "Arrival Delays", x = "Delay (minutes)", y = "Count") +
  theme_minimal()
```

# Relationship Between Departure and Arrival Delays

```{r}
#| label: plot-bivariate
#| fig-cap: "Relationship Between Departure and Arrival Delays"
ggplot(nycflights13::flights, aes(x = dep_delay, y = arr_delay)) +
  geom_point(alpha = 0.2, color = "purple") +
  xlim(-50, 300) + ylim(-50, 300) +
  labs(title = "Departure vs Arrival Delay", x = "Departure Delay (min)", y = "Arrival Delay (min)") +
  theme_minimal()
```

# Summary & Insights

In this analysis, I explored how departure delays relate to arrival delays for flights leaving NYC in 2013. First, I visualized the distribution of each variable separately. Both `dep_delay` and `arr_delay` are heavily skewed right, with most flights having small delays or even early arrivals. A smaller number of flights experienced long delays, which are clearly visible in the histogram tails.

In the scatter plot of departure delay vs arrival delay, we see a clear positive relationship. This means flights that depart late often arrive late as well—though not perfectly. The variation shows that some flights may make up lost time, while others may face additional delays. I chose histograms for univariate plots because they clearly show skewness and density, and a scatter plot for the bivariate relationship to visualize spread and correlation between the two variables.
